The challenge and promise of software citation for credit, identification, discovery, and reuse

### 2016-01-14 SC-challenge.tex

Modern science and engineering depend on software. 
To acknowledge traditional research results published in archival journals and conferences, we use citation, but no such accepted standard exists to credit the considerable efforts that go into software.
One method to increase the amount of software developed and shared is to
treat software as a publication.  And since research results now depend on the specific software used (e.g., version), proper citation---and the associated preservation---is necessary to ensure reproducibility.
\textbf{The main challenge of software citation is is that we need a way to uniquely identify released (published) software so that it can be cited by others who use it.}  We also need a process to curate and review software.

@@

Key challenges are the need to:
\begin{itemize}
\item standardize proper formats for citing software in publications,
\item identify necessary metadata associated with software for citation, 
\item establish mechanisms for software to cite other software (i.e., dependencies),
\item develop infrastructure to support indexing of software citations along with (or complementary to) the existing publication citation ecosystem, and
\item determine standard practices for peer review of software.
\end{itemize}
Although individual citation formats will vary based on the particular style of journals, conferences, or professional societies, citation of software in publications must minimally include software name, primary authors\slash contributors, DOI or other unique identifier, and location where software can be found (DOI or URL). 
The metadata needed for software to be cited includes software name, DOI, authors\slash contributors (with names and ORCIDs), software dependencies (in the form of DOIs), and any other people\slash artifacts that would be cited or acknowledged in a paper. \note{not sure about the previous two sentences?  explain more?  combine?}
This information should be provided in a CITATION file, potentially in JSON or XML formats (rather than plaintext) to allow automatic processing\slash parsing. \note{add DOAP?}

>>>>

Modern science and engineering depend on software. 
To acknowledge traditional research results published in archival journals and conferences, we use citation, but no such accepted standard exists to credit the considerable efforts that go into software.
One method to increase the amount of software developed and shared is to
treat a software release as a publication.  And since research results now depend on the specific software used (e.g., version), proper citation---and the associated preservation---is necessary to ensure reproducibility.
\textbf{The main challenge of software citation is is that we need a way to uniquely identify released (published) software so that it can be cited by others who use it.}  We also need a process to curate and review software.

@@

Key challenges are the need to:
\begin{itemize}
\item \note{added this, not sure if it really goes here} increase cultural acceptance of the concept of software as a digital product
\item standardize proper formats for citing software in publications,
\item identify necessary metadata associated with software for citation, 
\item establish mechanisms for software to cite other software (i.e., dependencies),
\item develop infrastructure to support indexing of software citations along with (or complementary to) the existing publication citation ecosystem, and
\item determine standard practices for peer review of software.
\end{itemize}
Although individual citation formats will vary based on the particular style of journals, conferences, or professional societies, citation of software in publications must minimally include software name, primary authors\slash contributors, DOI or other unique identifier, and location where the software has been published\slash archived (DOI or URL). 
The metadata needed for software to be cited includes software name, DOI, authors\slash contributors (with names and ORCIDs), software dependencies (in the form of DOIs), and any other people\slash artifacts that would be cited or acknowledged in a paper. \note{not sure about the previous two sentences?  explain more?  combine?}
This information should be provided in a CITATION file, potentially in JSON or XML formats (rather than plaintext) to allow automatic processing\slash parsing. \note{add DOAP?}
